---
# Heading1
author: Hugo Calligaro
title: "Sleep Multi Analysis"
output: html_notebook
last update: 2024-01-03
---

Requested packages:

```{r}
if(!require(openxlsx, quietly = TRUE)){install.packages("openxlsx")}
if(!require(ggplot2, quietly = TRUE)){install.packages("ggplot2")}
if(!require(tidyr, quietly = TRUE)){install.packages("tidyr")}
if(!require(tidyverse, quietly = TRUE)){install.packages("tidyverse")}
if(!require(matrixStats, quietly = TRUE)){install.packages("matrixStats")}
if(!require(rlang, quietly = TRUE)){install.packages("rlang")}
if(!require(emmeans, quietly = TRUE)){install.packages("emmeans")}
rm(list = ls())
```

Determine which animals need to be analysed for APP and WT (just comment those to exclude)

//!\\ Make sure the coma are right if you exclude some animals. Errors will show if you don't.

```{r}

  APP = c(
    "2504",
    "2505",
    "3160",
    "3161"
    )
  WT = c(
    "2670",
    "2671",
    "3086",
    "3159"
  )

```

Choose the criteria to open files:

```{r}
crit = c("Exp")

```

Importation of the list of files matching the criteria (crit2)

```{r}
  crit2 = c()
  crit3 = c()
  for (b in crit){
    if (is_empty(crit2)){
      crit2 = paste0(b)
      crit3 = paste0(b)
    } else {crit2=paste0(crit2,".+",b)
            crit3=paste0(crit3,"_",b)
    }
  }
  
  #setwd("C:/Users/hcall/OneDrive/Documents/R/Sleep")
  base_folder = getwd()
  setwd("output")
  a_unique = unique(list.files(pattern = crit2, recursive = T, ignore.case = T))
  setwd(base_folder)
  file_path = paste0("output/",a_unique)
  file_path = file_path[grep("~", file_path, invert = T)]
  files_pathway = file_path
  list_of_files_df = data.frame(file_name = c())
  sheet_list = getSheetNames(files_pathway[1])
  print(sheet_list)
```

Choose the sheet to open:

```{r}
sheet_to_open =   "Pre_wake"
```

Open all xlsx files and open the corresponding sheet as data.frame

```{r}
list_of_files_df = data.frame()
 for (a in 1:length(files_pathway)){
    list_of_files = unlist(data.table::tstrsplit(a_unique[a],"/", fill=NA, type.convert=FALSE, names=F))[2]
    list_of_files_data = paste0("Data_",list_of_files)
    list_of_files_data = gsub(".xlsx","", list_of_files_data)
    list_of_files_df = rbind(list_of_files_df, list_of_files_data)
    temp_ = read.xlsx(files_pathway[a],sheet = sheet_to_open)
    if (names(temp_)[1]== "ZT"){
      temp_ = arrange(temp_, ZT)}
    names_col = names(temp_)
    splits_temp=unlist(data.table::tstrsplit(list_of_files,"_", fill=NA, type.convert=FALSE, names=F))
    new_names_col = paste0(names_col,"_",splits_temp[2],"_",splits_temp[1])
    names(temp_)=new_names_col
    assign(list_of_files_data,temp_)
  }
  names(list_of_files_df)= "file_name"
  
head(temp_)

```

List of column in selected sheet:

```{r}

print(names_col)
```

Choose the column to combine:

```{r}
para_= "total_sleep_measured"

if (is_empty(which(names_col == para_))){
  print("///// Wrong column name, just copy-paste, please")
        #, you know you can't type properly \\\\\\")
}

if (!is_empty(grep("Temp_Mean", para_)) | !is_empty(grep("Activity_Mean", para_)) | !is_empty(grep("Temperature", para_)) | !is_empty(grep("Activity", para_))){
    APP = c("2504","2505","3160","3161")
    WT = c("2670","2671","3086","3159")
} else {
  if (!is_empty(grep("Exp", crit2))){
      APP = c("2504","2505","3160","3161")
      WT = c("2670","2671","3086")
  } else if (!is_empty(grep("Green", crit2))){
      APP = c("2504","3160","3161")
      WT = c("2671","3086")
  } else if (!is_empty(grep("Blue", crit2))){
      APP = c("2504","3160","3161")
      WT = c("2671","3086")
  }
}

```

Combine the column selected with para\_

```{r}
  ###Combine DF
  #ZT = seq(0,22, by=2)
  x_axis = temp_[1]  
  combine_df = data.frame(x_axis = x_axis)
  names(combine_df)="x_axis"
  col_added_name_2 = c()
  parameter_to_analyse = which(names_col == para_)
  for (i in 1:nrow(list_of_files_df)) {
    # Get the current data frame name
    df_name <- list_of_files_df$file_name[i]
    
    # Access the data frame by name
    current_df <- get(df_name)
    col_to_add = current_df[parameter_to_analyse]
    combine_df = cbind(combine_df,col_to_add)
    
    #Prefix extraction
    col_added_name = unlist(data.table::tstrsplit(names(col_to_add),"_",fill=NA, type.convert=FALSE,names=F))
    ID_to_del = paste0("_",col_added_name[length(col_added_name)])
    col_added_name_2[i] = gsub(ID_to_del,"", names(col_to_add))
  }
head(combine_df)
```

Calculate the average and SD, then save as xlsx

```{r}

  if (length(APP)>1){
    APP_comb = c(APP[1])
    for (a in 2:length(APP)){
      APP_comb=paste0(APP_comb,"|",APP[a])
    }
    combine_df$APP_mean = rowMeans(combine_df[, grep(APP_comb, names(combine_df))])
    combine_df$APP_SD = rowSds(as.matrix(combine_df[, grep(APP_comb, names(combine_df))]))
    combine_df$APP_SEM = (combine_df$APP_SD / sqrt(length(APP)))
  } else {
    APP_comb = APP
    combine_df$APP_mean = combine_df[, grep(APP_comb, names(combine_df))]
    combine_df$APP_SD = rep(NA,nrow(combine_df))
    combine_df$APP_SEM = rep(NA,nrow(combine_df))
  }

 if (length(WT)>1){
    WT_comb = c(WT[1])
    for (a in 2:length(WT)){
      WT_comb=paste0(WT_comb,"|",WT[a])
    }
    combine_df$WT_mean = rowMeans(combine_df[, grep(WT_comb, names(combine_df))])
    combine_df$WT_SD = rowSds(as.matrix(combine_df[, grep(WT_comb, names(combine_df))]))
    combine_df$WT_SEM = (combine_df$WT_SD / sqrt(length(WT)))
  } else {
    WT_comb = WT
    combine_df$WT_mean = combine_df[, grep(WT_comb, names(combine_df))]
    combine_df$WT_SD = rep(NA,nrow(combine_df))
    combine_df$WT_SEM = rep(NA,nrow(combine_df))
  }
head(combine_df[,(length(combine_df)-5):(length(combine_df))])

if (dir.exists("combine_df/")==F){
  dir.create("combine_df/")
}
base_folder_ = paste0("combine_df/",crit[1])
if (dir.exists(base_folder_)==F){
  dir.create(base_folder_)
}

combine_df_exp_name = paste0("combine_df/",crit[1],"/",para_)
if (dir.exists(combine_df_exp_name)==F){
  dir.create(combine_df_exp_name)
}

head(temp_)

testwb = createWorkbook()
addWorksheet(testwb, para_)
writeData(testwb,para_, x=combine_df)
filename = paste0(combine_df_exp_name,"/",para_,"_",crit3,".xlsx")
saveWorkbook(testwb, file = filename, overwrite = T)

comb_unique = unique(list.files(patter = para_, recursive = T, ignore.case = T))
comb_unique = comb_unique[grep(crit[1],comb_unique)]
comb_unique = comb_unique[grep("_all",comb_unique,invert=T)]
comb_unique = comb_unique[grep(".xlsx",comb_unique)]

combine_workbook = createWorkbook()
 for (a in 1:length(comb_unique)){
    comb_unique_temp = unlist(data.table::tstrsplit(comb_unique[a],"/", fill=NA, type.convert=FALSE, names=F))
    comb_unique_temp2 = comb_unique_temp[length(comb_unique_temp)]
    comb_unique_temp2 = gsub(".xlsx","", comb_unique_temp2)
    list_of_files_df = rbind(list_of_files_df, list_of_files_data)
    temp_ = read.xlsx(comb_unique[a])
    assign(comb_unique_temp2,temp_)
    addWorksheet(combine_workbook, comb_unique_temp2)
    writeData(combine_workbook, comb_unique_temp2, x= temp_)
 }
# 
# filename_combine_wb = paste0(combine_df_exp_name,"/",para_,"_all.xlsx")
# saveWorkbook(combine_workbook, file = filename_combine_wb, overwrite = T)

```

Create GraphPlot

```{r}
png_exp_name = paste0("png/",crit3)
if (dir.exists("png/")==F){
  dir.create("png/")
}
if (dir.exists(png_exp_name)==F){
  dir.create(png_exp_name)
}
png_name = paste0(png_exp_name,"/",names(col_to_add),".svg")
svg(filename = png_name,         # File name
    width = 8, height = 5
    #res = 300
    )
g = c()
if (is.numeric(combine_df$x_axis)){
  g = ggplot(combine_df, aes(x=x_axis))+

    geom_errorbar(aes(ymin = APP_mean, ymax = APP_mean+APP_SEM), width = 0.25)+
    geom_errorbar(aes(ymin = WT_mean-APP_SEM, ymax = WT_mean), width = 0.25)+
    geom_line(aes(y = APP_mean, color="APP"),linewidth = 1)+
    geom_line(aes(y=WT_mean, color="WT"),linewidth = 1)+
    geom_point(aes(y =APP_mean, color="APP"), size=2)+
    geom_point(aes(y=WT_mean, color="WT"), size = 2)+
    labs(y=names(col_to_add), title = crit2, colour = "Genotype")+
    geom_vline(xintercept = 12, linetype = "dotted")+
    scale_x_continuous(limits = c(-1,24), n.breaks=14)+
    # scale_y_continuous(limits = c(35,38))+
    scale_color_manual(values = c(APP = "#D55E00", WT = "#009E73"), 
                     labels = c( WT = "WT",APP = "APP")) +
    #geom_vline(xintercept = c(14,22), linetype ="dashed")+
    #geom_vline(xintercept = c(16,18))+
    ylim(0,NA)+
    theme_classic()
      if (crit2 == "TRF.+Chronic"){
      g = g + geom_vline(xintercept = c(14,22), linetype ="dashed")
      } else if (crit2 == "TRF.+Light"){
                g = g + geom_vline(xintercept = c(14,22), linetype ="dashed")+
                geom_vline(xintercept = c(16,18))
      } else if (crit2 == "TRF.+Pulse2"){
                g = g + geom_vline(xintercept = c(16,18))
      }
} else if (is.character(combine_df$x_axis)){
    if (combine_df$x_axis[1] == "0-.5"){
    #combine_df$x_axis = fct_inorder(combine_df$x_axis)
    ### Create data2 here
    combine_df = combine_df[1:6,]
    Genotype = c(rep("APP",6), rep("WT",6))
    comp_ = rep(c("0-.5", "0.5-1", "1-2", "2-4", "4-8", "8-16"),2)
    Var_ = c(combine_df$APP_mean,combine_df$WT_mean)
    SEM_ = c(combine_df$APP_SEM,combine_df$WT_SEM)
    data2 = data.frame(Genotype, comp_, Var_, SEM_)
    data2$Genotype <- factor(data2$Genotype, levels = c("WT", "APP"))
    data2$comp_ = factor(data2$comp_, levels = c("0-.5", "0.5-1", "1-2", "2-4", "4-8", "8-16"))
    combine_df$x_axis_continuous = seq(1,nrow(combine_df),by = 1)
    custom_labels <- c("0-.5", "0.5-1", "1-2", "2-4", "4-8", "8-16")
    
    g = ggplot(data2, aes(x=comp_, y=Var_, fill = Genotype))+
      geom_bar(position = "dodge", stat = "identity") +
      geom_errorbar(aes(ymin = Var_ - SEM_, ymax = Var_ + SEM_), stat = "identity",
                    position = position_dodge(width = 0.9), colour = "black", width = 0.2) +
      geom_hline(yintercept = 0) +
      labs(y=names(col_to_add), x ="Length of uninturrepted bouts of sleep (min)", title = crit2, colour = "Genotype")+
      scale_fill_manual(values = c(APP = "#D55E00", WT = "#009E73"))+
      theme_classic()
      
      
      
      # geom_line(aes(y = APP_mean, color="APP"),linewidth = 1)+
      # geom_errorbar(aes(ymin = APP_mean-APP_SEM, ymax = APP_mean+APP_SEM), width = 0.25)+
      # geom_point(aes(y =APP_mean), size=3, shape = 3)+
      # geom_line(aes(y=WT_mean, color="WT"),linewidth = 1)+
      # geom_errorbar(aes(ymin = WT_mean-WT_SEM, ymax = WT_mean+WT_SEM), width = 0.25)+
      # geom_point(aes(y=WT_mean), size = 3, shape = 3)+
      
      
      
    } else if (combine_df$x_axis[1] == "Light" & length(combine_df$x_axis) == 8){
    Genotype = c(rep("APP",8), rep("WT",8))
    Phase = rep(c(rep("Light",4), rep("Dark",4)),2)
    Stage = rep(c("Wake","NREM","REM","Total Sleep"),4)
    Var_ = c(combine_df$APP_mean,combine_df$WT_mean)
    SEM_ = c(combine_df$APP_SEM,combine_df$WT_SEM)
    data2 = data.frame(Genotype, Phase, Stage, Var_, SEM_)
    data2$Phase <- factor(data2$Phase, levels = c("Light", "Dark"))
    data2$Genotype <- factor(data2$Genotype, levels = c("WT", "APP"))
    data2$Stage = factor(data2$Stage, levels = c("Wake","NREM","REM","Total Sleep"))
    
    
    g = ggplot(data2, aes(x = Stage, y = Var_, fill = Genotype)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_errorbar(aes(ymin = Var_ - SEM_, ymax = Var_ + SEM_), stat = "identity",
                    position = position_dodge(width = 0.9), colour = "black", width = 0.2) +
      geom_hline(yintercept = 0) +
      labs(x=NULL, y=names(col_to_add), title = crit2, colour = "Genotype")+
      theme_classic() +
      scale_fill_manual(values = c(APP = "#D55E00", WT = "#009E73")) +  # Adjust colors as needed
      facet_wrap(~Phase, scales = "fixed", nrow = 1)
    
    } else if (combine_df$x_axis[1] == "Light" & length(combine_df$x_axis) == 12){
    data2=c()
    Genotype = c(rep("APP",12), rep("WT",12))
    Phase = rep(c(rep("Light",6), rep("Dark",6)),2)
    Cluster = rep(c("Cluster1","Cluster2","Cluster3","Cluster4","Cluster5","Cluster6"),4)
    Var_ = c(combine_df$APP_mean,combine_df$WT_mean)
    SEM_ = c(combine_df$APP_SEM,combine_df$WT_SEM)
    data2 = data.frame(Genotype, Phase, Cluster, Var_, SEM_)
    data2$Phase <- factor(data2$Phase, levels = c("Light", "Dark"))
    data2$Genotype <- factor(data2$Genotype, levels = c("WT", "APP"))
    data2$Cluster = factor(data2$Cluster, levels = c("Cluster1","Cluster2","Cluster3","Cluster4","Cluster5","Cluster6"))
    
      g = ggplot(data2, aes(x = Cluster, y = Var_, fill = Genotype)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_errorbar(aes(ymin = Var_ - SEM_, ymax = Var_ + SEM_), stat = "identity",
                    position = position_dodge(width = 0.9), colour = "black", width = 0.2) +
      geom_hline(yintercept = 0) +
      labs(x=NULL, y=names(col_to_add), title = crit2, colour = "Genotype")+
      theme_classic() +
      scale_fill_manual(values = c(APP = "#D55E00", WT = "#009E73")) +  # Adjust colors as needed
      facet_wrap(~Phase, scales = "fixed", nrow = 1)
    
    } else if (combine_df$x_axis[1] == "Light" & length(combine_df$x_axis) == 2){
          data2=c()
    Genotype = c(rep("APP",2), rep("WT",2))
    Phase = rep(c("Light","Dark"),2)
    #Cluster = rep(c("Cluster1","Cluster2","Cluster3","Cluster4","Cluster5","Cluster6"),4)
    Var_ = c(combine_df$APP_mean,combine_df$WT_mean)
    SEM_ = c(combine_df$APP_SEM,combine_df$WT_SEM)
    data2 = data.frame(Genotype, Phase, Var_, SEM_)
    data2$Phase <- factor(data2$Phase, levels = c("Light", "Dark"))
    data2$Genotype <- factor(data2$Genotype, levels = c("WT", "APP"))
    #data2$Cluster = factor(data2$Cluster, levels = c("Cluster1","Cluster2","Cluster3","Cluster4","Cluster5","Cluster6"))
    
      g = ggplot(data2, aes(x = Phase, y = Var_, fill = Genotype)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_errorbar(aes(ymin = Var_ - SEM_, ymax = Var_ + SEM_), stat = "identity",
                    position = position_dodge(width = 0.9), colour = "black", width = 0.2) +
      geom_hline(yintercept = 0) +
      labs(x=NULL, y=names(col_to_add), title = crit2, colour = "Genotype")+
      theme_classic() +
      scale_fill_manual(values = c(APP = "#D55E00", WT = "#009E73"))  # Adjust colors as needed
      
    } else {print("Type of graph not supported yet")}
} else {print("Unrecognized x-axis")}
g
dev.off() 

g
```

Combine APP and WT as same column for stat analysis

```{r}
if (is.numeric(combine_df$x_axis)){
ZT2 = seq(min(combine_df$x_axis), max(combine_df$x_axis), by = (combine_df$x_axis[2]-combine_df$x_axis[1]))
combine_APP = combine_df[, grep(APP_comb, names(combine_df))]
  ALLAPP=c()
  for (a in 1:length(combine_APP)){
    ALLAPP = c(ALLAPP,combine_APP[,a])
  }
  all_APP_df = data.frame(ZT=as.factor(rep(ZT2, length(combine_APP))),Geno=rep("APP", length(combine_APP)*nrow(combine_APP)), Param = ALLAPP)
  
  combine_WT = combine_df[, grep(WT_comb, names(combine_df))]
  ALLWT=c()
  if (length(WT)>1){
    for (a in 1:length(combine_WT)){
      ALLWT = c(ALLWT,combine_WT[,a])
    }
    all_WT_df = data.frame(ZT=as.factor(rep(ZT2, length(combine_WT))),Geno=rep("WT", length(combine_WT)*nrow(combine_WT)), Param = ALLWT)
    
  } else {
    ALLWT = combine_WT
    all_WT_df = data.frame(ZT = ZT2, Geno = rep("WT",length(ZT2)), Param =ALLWT)
  }
} else if (is.numeric(combine_df$x_axis_continuous)){
  x_axis = combine_df$x_axis_continuous
  
  combine_APP = combine_df[, grep(APP_comb, names(combine_df))]
  ALLAPP=c()
  for (a in 1:length(combine_APP)){
    ALLAPP = c(ALLAPP,combine_APP[,a])
  }
  all_APP_df = data.frame(comp_=as.factor(rep(x_axis, length(combine_APP))),Geno=rep("APP", length(combine_APP)*nrow(combine_APP)),
                          Param = ALLAPP)
  
  combine_WT = combine_df[, grep(WT_comb, names(combine_df))]
  ALLWT=c()
  for (a in 1:length(combine_WT)){
      ALLWT = c(ALLWT,combine_WT[,a])
    }
    all_WT_df = data.frame(comp_=as.factor(rep(x_axis, length(combine_WT))),Geno=rep("WT", length(combine_WT)*nrow(combine_WT)), Param = ALLWT)
  
  
} else if (!is.numeric(combine_df$x_axis) & length(combine_df$x_axis) == 8){
  x_axis = combine_df$x_axis
  
  combine_APP = combine_df[, grep(APP_comb, names(combine_df))]
  ALLAPP=c()
  for (a in 1:length(combine_APP)){
    ALLAPP = c(ALLAPP,combine_APP[,a])
  }
  all_APP_df = data.frame(comp_=as.factor(rep(x_axis, length(combine_APP))),Geno=rep("APP", length(combine_APP)*nrow(combine_APP)),
                          Param = ALLAPP)
  
  combine_WT = combine_df[, grep(WT_comb, names(combine_df))]
  ALLWT=c()
  for (a in 1:length(combine_WT)){
      ALLWT = c(ALLWT,combine_WT[,a])
    }
    all_WT_df = data.frame(comp_=as.factor(rep(x_axis, length(combine_WT))),Geno=rep("WT", length(combine_WT)*nrow(combine_WT)), Param = ALLWT)
  
  All_APP_WT = rbind(all_APP_df,all_WT_df)
  nb_sample = nrow(All_APP_WT)/4
  stage_sleep = rep(c("Wake","NREM","REM","Total Sleep"), nb_sample)
  All_APP_WT = cbind(All_APP_WT, stage_sleep)
  
  All_APP_WT_Light = subset(All_APP_WT, comp_ == "Light")
  All_APP_WT_Dark = subset(All_APP_WT, comp_ == "Dark") 

} else {print("Not implemented yet")}
```

Two way analysis

```{r}
print(paste0(crit2," ",colnames(col_to_add)))
 if (combine_df$x_axis[1]==0){
  All_APP_WT = rbind(all_APP_df,all_WT_df)
  res.aov2 = aov(Param~ZT*Geno, data=All_APP_WT)
  summary(res.aov2)
 } else if (!is.numeric(combine_df$x_axis) & length(combine_df$x_axis) == 8){
  print("Light phase:")
   res.aov_light = aov(Param~stage_sleep*Geno, data=All_APP_WT_Light)
  print(summary(res.aov_light))
  print(" ")
  print("Dark phase:")
   res.aov_dark = aov(Param~stage_sleep*Geno, data=All_APP_WT_Dark)
  print(summary(res.aov_dark))
 } else if (is.numeric(combine_df$x_axis_continuous) & combine_df$x_axis_continuous[1]==1){ ### find better way to discriminate it
  All_APP_WT = rbind(all_APP_df,all_WT_df)
  res.aov2 = aov(Param~comp_*Geno, data=All_APP_WT)
  summary(res.aov2)
   
   } else {print("Not implemented yet")}
```

```{r}
print(paste0(crit2," ",colnames(col_to_add)))
 if (is.numeric(combine_df$x_axis)){
  emmeans_results <- emmeans(res.aov2, ~ Geno | ZT)
  pairwise_results <- contrast(emmeans_results, method = "pairwise")
  summary(pairwise_results)
 } else if (!is.numeric(combine_df$x_axis) & length(combine_df$x_axis) == 8){
   print("Light phase:")
  emmeans_results <- emmeans(res.aov_light, ~ Geno | stage_sleep)
  pairwise_results <- contrast(emmeans_results, method = "pairwise")
  print(summary(pairwise_results))
  print(" ")
  print("Dark phase:")
  emmeans_results <- emmeans(res.aov_dark, ~ Geno | stage_sleep)
  pairwise_results <- contrast(emmeans_results, method = "pairwise")
  print(summary(pairwise_results))
 } else if (is.numeric(combine_df$x_axis_continuous)){
  emmeans_results <- emmeans(res.aov2, ~ Geno | comp_)
  pairwise_results <- contrast(emmeans_results, method = "pairwise")
  summary(pairwise_results) 
 } else {print("Not implemented yet")}
```
